--- 
title: "Section 1.1 TB incidence" 
author: "Mathieu Bastard, reformatted/refactored by Hazim Timimi" 
date: "`r Sys.Date()`" 
knit: (function(inputFile, encoding) {
      out_dir <- "./local";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
output:
  html_fragment:
    # Don’t include a table of contents
    toc: no
    # Set standard figure width to 12 inches
    fig_width: 12
    # Don’t write figure captions
    fig_caption: FALSE
    # Don't embed external resources (stylesheets, JS libraries) in the output
    self_contained: FALSE
    # Don't include <div> tags around a header and the content found until the next header
    section_divs: FALSE  

# To run this file and store output as html:
# rmarkdown::render(here::here("report/ch1-1.rmd"), output_file = "ch1-1.html", output_dir = here::here("report/local/"))
--- 


```{r setup, include=FALSE} 
# Chapter 1
# Set options, load data, utility functions 

knitr::opts_chunk$set(echo = FALSE, 
                      results = "asis",
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE)  # TEMP error=TRUE for debugging!

# Kill any attempt at using factors, unless we explicitly want them!
options(stringsAsFactors=FALSE)
options(scipen=999)


# Load output packages ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
library(data.table)
library(gtbreport)
library(here)
library(kableExtra)
library(gridExtra)
library(whomap)
library(ggrepel)
library(ggpubr)
library(RColorBrewer)
library(tidyverse)
library(stringr)
library(jsonlite)
library(dplyr)



# Prepare data for the chapter
source(here('report/ch1_prepare_data.R'))

# Load R functions ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

source(here("report/functions/html_links.R"))



# Set output options ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 


# Show static chart in addition to Kendo chart?
show_static = FALSE

# Save underlying data files as CSV and charts as PDF files?
pdf_csv_folder = here::here("report/local/figures/ch1-1")
save_csv = TRUE
save_pdf = TRUE

# Create the output folder (only if it doesn't yet exist)
dir.create(pdf_csv_folder, showWarnings = FALSE, recursive = TRUE)


``` 

```{r css_js}
# Add standard stylesheets and javascript to support kendo
cat(writeLines(readLines(here("report/resources/headers.htm"))))
```

```{css, echo=FALSE}

/* Table 1.1.1 */
/* Recreating simple striped bootstrap table */
#burden_num_table {
  border-spacing: 0;
  border-collapse: collapse;
  margin-top: 1em;
  margin-bottom: 1em;
  /* Next two lines to allow horizontal scrolling on narrow screens */
  display: block;
  overflow-x: auto;
}

#burden_num_table th {
  border-bottom: 2px solid #DDDDDD;
  padding: 8px;
}

#burden_num_table td {
  border-top: 1px solid #DDDDDD;
  padding: 8px;
}

/* light gray for odd rows */
#burden_num_table tr:nth-child(odd) td {
  background-color: #F5F5F5;	
}

/* Bold for the final row with thick line above */
#burden_num_table tr:last-child td {
  border-top: 2px solid #DDDDDD;
  font-weight:bold;	
}

/* light gray when hovering over a row */
#burden_num_table tr:hover td {
  background-color: #DDDDDD;
}

/* Centre-align all column headings except for the first */
#burden_num_table th:not(:first-child) {
  text-align: center !important;
}

/* prevent numbers from wrapping in any of the columns */
#burden_num_table td {
  white-space: nowrap;
}


```

```{r}
# to format 100 000 as a R object, avoid breaking lines
cm=formatC(1e5, format="f", big.mark=" ", digits=0)

# Estimates year
yr=2023
```



## 1.1 TB incidence

_Draft! Prepared `r Sys.Date()` using country-reported data snapshot files from `r snapshot_date` and incidence estimates from `r est_date`_

`r lnk("Box 1.1.1")` provides a summary of the methods used to produce estimates of TB incidence for the period 2010–`r yr`. For 47 countries that experienced considerable disruptions to TB services during the COVID-19 pandemic (in terms of reductions in TB case notifications that were a clear departure from historic trends), these methods include the use of country or region-specific dynamic models for the period 2020–2023. Although the numbers of people being newly diagnosed and notified as a TB case recovered to pre-pandemic levels in almost all of these countries by 2022 or 2023 (<span class="red">Section 2.1</span>), models have continued to be used because the effects of even temporary disruptions on incidence can persist for several years, as a result of the slow natural history of TB (`r ref_lnk("1, 2, 3")`).

As new data become available to inform estimates of TB incidence from 2024 onwards, the use of these models can be phased out.


An estimated global total of `r round(global$inc.num[global$year==2023]/1e6,1)` million people (95% uncertainty interval [UI]: `r round(global$inc.lo.num[global$year==2023]/1e6,1)`–`r round(global$inc.hi.num[global$year==2023]/1e6,1)` million) fell ill with TB in `r yr`, equivalent to `r round(global$inc[global$year==2023],0)` incident cases (95% UI: `r round(global$inc.lo[global$year==2023],0)`–`r round(global$inc.hi[global$year==2023],0)`) per `r cm` population (`r lnk("Table 1.1.1")`). Among all incident TB cases, `r round(100*global$inc.h.num[global$year==2023]/global$inc.num[global$year==2023],1)`% were among people living with HIV. Most TB cases in 2023 were in the WHO regions of South-East Asia (`r round(100*regional$inc.num[regional$year==2023 & regional$g.whoregion=="SEA"]/global$inc.num[global$year==2023],0)`%), Africa (`r round(100*regional$inc.num[regional$year==2023 & regional$g.whoregion=="AFR"]/global$inc.num[global$year==2023],0)`%) and the Western Pacific (`r round(100*regional$inc.num[regional$year==2023 & regional$g.whoregion=="WPR"]/global$inc.num[global$year==2023],0)`%), with smaller shares in the Eastern Mediterranean (`r round(100*regional$inc.num[regional$year==2023 & regional$g.whoregion=="EMR"]/global$inc.num[global$year==2023],1)`%), the Americas (`r round(100*regional$inc.num[regional$year==2023 & regional$g.whoregion=="AMR"]/global$inc.num[global$year==2023],1)`%) and Europe (`r round(100*regional$inc.num[regional$year==2023 & regional$g.whoregion=="EUR"]/global$inc.num[global$year==2023],1)`%). 


### `r anch("Table 1.1.1")`<span class="red">Table 1.1.1</span> Global and regional estimates of TB incidence, numbers (in thousands) and rates (per `r cm` population) in `r yr`
<div class="subhead">Low and high are the 5th and 95th percentiles of the uncertainty interval (UI).</div>
```{r tab_1.1.1}

tab.header <- c('Region or country group','Population','Best estimate','Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High'
                  )

knitr::kable(cbind(tabinc[c(32:37, 31, 38), 1:8], tabmort[c(32:37, 31, 38), 2:7]),
             format = "html",
             align = 'rccccccccccccc',
             col.names = tab.header,
             # Add a table ID so that it can be styled using extra CSS in Sitefinity
             table.attr = "id='burden_num_table'") |>
add_header_above(header = c(" " = 2, "Total" = 3, "People with HIV" = 3, "Total" = 3, "People with HIV" = 3)) |>
add_header_above(header = c(" " = 2, "Number of cases (in thousands)" = 6, "Rate per 100 000 population" = 6)) 
  # pack_rows("High burden countries", 1, 30) |>
  # pack_rows('Groupings', 31, 38) 


```

<hr />
<br />

```{r}
inc.cases.23=round(global$inc.num[global$year==2023]/1e6,1)
inc.cases.23.lo=round(global$inc.lo.num[global$year==2023]/1e6,1)
inc.cases.23.hi=round(global$inc.hi.num[global$year==2023]/1e6,1)


inc.cases.22=round(global$inc.num[global$year==2022]/1e6,1)
inc.cases.22.lo=round(global$inc.lo.num[global$year==2022]/1e6,1)
inc.cases.22.hi=round(global$inc.hi.num[global$year==2022]/1e6,1)

inc.cases.21=round(global$inc.num[global$year==2021]/1e6,1)
inc.cases.21.lo=round(global$inc.lo.num[global$year==2021]/1e6,1)
inc.cases.21.hi=round(global$inc.hi.num[global$year==2021]/1e6,1)

inc.cases.20=round(global$inc.num[global$year==2020]/1e6,1)
inc.cases.20.lo=round(global$inc.lo.num[global$year==2020]/1e6,1)
inc.cases.20.hi=round(global$inc.hi.num[global$year==2020]/1e6,1)

```


The estimated `r inc.cases.23` million people (95% UI: `r inc.cases.23.lo`–`r inc.cases.23.hi` million) who fell ill with TB worldwide in 2023 is an increase from `r inc.cases.22` million in 2022 (95% UI: `r inc.cases.22.lo`.0–`r inc.cases.22.hi` million), `r format(inc.cases.21,nsmall=1)` million (95% UI: `r inc.cases.21.lo`–`r inc.cases.21.hi` million) in 2021 and `r format(inc.cases.20,nsmall=1)` million (95% UI: `r inc.cases.20.lo`–`r inc.cases.20.hi` million) in 2020, continuing the reversal of the downward trend that had been sustained for many years up to 2020 (`r lnk("Fig. 1.1.1")`, panel a). The TB incidence rate (new cases per `r cm` population per year) is estimated to have increased only slightly (by `r round((100*global$inc[global$year==yr]/global$inc[global$year==yr-1])-100,1)`%) between 2022 and 2023 (`r lnk("Fig. 1.1.1")`, panel b), a much smaller amount than the `r round((100*global$inc[global$year==yr-1]/global$inc[global$year==yr-2])-100,1)`% between 2021 and 2022.

The three consecutive years of increasing global TB incidence from 2021–2023 mean that in 2023, the TB incidence rate returned to the level of 2018  (`r lnk("Fig. 1.1.1")`, panel b). Globally, the net reduction in the TB incidence rate from 2015 to 2023 was `r 100-round(100*global$inc[global$year==yr]/global$inc[global$year==2015],1)`%, far from the WHO End TB Strategy milestone of a `r (1-inc.milestone.vs.2015)*100`% reduction by `r inc.mort.milestone.yr` (`r ref_lnk("4, 5")`).

### `r anch("Fig. 1.1.1")`<span class="red">Fig. 1.1.1</span> Global trends in the estimated number of incident TB cases (a) and the incidence rate (b), 2010&#8211;`r yr`
<div class="subhead">Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1-inc.milestone.vs.2015)*100`% reduction in the TB incidence rate between 2015 and `r inc.mort.milestone.yr`.</div>

```{r}



f1.1.1a <- 
  qplot(
    year,
    inc.num / 1e6,
    data = subset(global,year>=2010),
    geom = 'line',
    colour = I('blue')
  ) +
  geom_ribbon(
    aes(year, ymin = inc.lo.num / 1e6, ymax = inc.hi.num / 1e6),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, inc.h.num / 1e6), colour = I('lightskyblue2')) +
  geom_line(aes(year, c.newinc / 1e6)) +
  geom_ribbon(
    aes(year, ymin = inc.h.lo.num / 1e6, ymax = inc.h.hi.num / 1e6),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  ylab('Millions per year') + xlab('') +
  annotate(
    'text',
    x = 2010,
    y = 7.9,
    label = 'Total',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2015,
    y = 4.2,
    label = 'Case notifications of people \nnewly diagnosed with TB',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2010,
    y = 0.9,
    label = 'People living with HIV',
    hjust = 0,
    size = I(4)
  ) +
  expand_limits(y = 1) +
  # scale_y_log10() +
  scale_x_continuous(breaks=seq(2010,2022,2)) +
  #scale_x_continuous(breaks=c(2010,2015,2019,2023)) +
  
  theme_bw(base_size = 14)+
  theme_gtb()


f1.1.1a_data=global[year>=2010,.(year,inc.num,inc.lo.num,inc.hi.num,inc.h.num,inc.h.lo.num,inc.h.hi.num,c.newinc)]

output_ggplot(f1.1.1a, f1.1.1a_data, show_static, pdf_csv_folder, save_csv, save_pdf)


## Fig1b

f1.1.1b <- 
  qplot(
    year,
    inc,
    data = subset(global,year>=2010),
    geom = 'line',
    colour = I('blue')
  ) +
  geom_ribbon(
    aes(year, ymin = inc.lo, ymax = inc.hi),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, inc.h), colour = I('lightskyblue2')) +
  geom_line(aes(year, newinc)) +
  geom_ribbon(
    aes(year, ymin = inc.h.lo, ymax = inc.h.hi),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  geom_hline(aes(yintercept = inc.milestone), linetype = I(2)) +
  
  ylab('Rate per 100 000 population per year') + xlab('') +
  annotate(
    'text',
    x = 2020,
    y = 60,
    label = paste(inc.mort.milestone.yr, 'milestone'),
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2012,
    y = 190,
    label = 'Total',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2010,
    y = 110,
    label = 'Case notifications of people \nnewly diagnosed with TB',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2010,
    y = 10,
    label = 'People living with HIV',
    hjust = 0,
    size = I(4)
  ) +  
  expand_limits(y = 0) +
  # scale_y_log10() +
  scale_x_continuous(breaks=seq(2010,2022,2)) +
  theme_bw(base_size = 14)+
  theme_gtb()


f1.1.1b_data=global[year>=2010,.(year,inc,inc.lo,inc.hi,inc.h,inc.h.lo,inc.h.hi,newinc)]

output_ggplot(f1.1.1b , f1.1.1b_data, show_static, pdf_csv_folder, save_csv, save_pdf)


```


<div class="col-md-6">
#### (a) Number
<div id="fig_1_1_1a"></div>
</div>
<div class="col-md-6">
#### (b) Rate per `r cm` population
<div id="fig_1_1_1b"></div>
</div>

<hr />
<br />

The estimated increases in TB incidence in 2021–2023 are the consequence of disruptions to TB diagnosis and treatment during the COVID-19 pandemic, when the reported number of people newly diagnosed with TB fell from `r round(global$c.newinc[global$year==2019]/1e6,1)` million in 2019 to `r round(global$c.newinc[global$year==2020]/1e6,1)` million in 2020 and `r round(global$c.newinc[global$year==2021]/1e6,1)` million in 2021 (<span class="red">Section 2.1</span>). These reductions are assumed to have resulted in an increase in the number of people with undiagnosed and untreated TB (with the dynamic models used for 47 countries also allowing for reductions in transmission during periods of lockdowns as well as some level of underreporting). 

The most immediate impact of an increase in the number of people with undiagnosed and untreated TB was an increase in TB mortality (<span class="red">Section 1.2</span>), but it also resulted in increased transmission of TB infection and then, with a lag time (given that the time from infection to progression to TB disease ranges among individuals from months to years), a subsequent increase in the number of people developing TB disease. Similarly, the most immediate impact of the big global recovery in the reported number of people newly diagnosed with TB in 2022 and 2023 (<span class="red">Section 2.1</span>) has been a reduction in TB mortality. Reversing the global increase in TB incidence will take longer. Based on current trends, the global increase in the number of incident cases is slowing and may stabilize or reverse in the next one or two years.

In `r yr`, `r f1.1.2_top |> nrow() |> int2word()` countries accounted for more than two thirds of global TB cases: `r f1.1.2_top[1,"country.in.text.EN"]` (`r ftb(f1.1.2_top[1,"pct"])`%), `r f1.1.2_top[2,"country.in.text.EN"]` (`r ftb(f1.1.2_top[2,"pct"])`%), `r f1.1.2_top[3,"country.in.text.EN"]` (`r ftb(f1.1.2_top[3,"pct"])`%), `r f1.1.2_top[4,"country.in.text.EN"]` (`r ftb(f1.1.2_top[4,"pct"])`%), `r f1.1.2_top[5,"country.in.text.EN"]` (`r ftb(f1.1.2_top[5,"pct"])`%), `r f1.1.2_top[6,"country.in.text.EN"]` (`r ftb(f1.1.2_top[6,"pct"])`%), `r f1.1.2_top[7,"country.in.text.EN"]` (`r ftb(f1.1.2_top[7,"pct"])`%) and `r f1.1.2_top[8,"country.in.text.EN"]` (`r ftb(f1.1.2_top[8,"pct"])`%) (`r lnk("Fig. 1.1.2")`). 


### `r anch("Fig. 1.1.2")`<span class="red">Fig. 1.1.2</span> Estimated number of incident TB cases in `r yr`, for countries with at least `r cm` incident cases

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1.2, fig.alt="Estimated number of incident TB cases in 2023, for countries with at least 100 000 incident cases"} 


dta <- est[year == yr & inc.num > 1e5, .(iso3, size = inc.num)]
dta <- merge(dta, cty[,.(iso3, country)], by='iso3')

f1.1.2_plot <-
bubblemap(dta, legend.pos = c(0.14, 0.5), legend.title = 'Number of\nincident cases',
          bubble.col = "#08519C",
          water.col = 'white',
          scale.limits = c(100000, 3000000),
          scale.breaks = c(100000, 500000,1000000,2000000),
          scale.labels = c('100 000', '500 000','1 000 000', '2 000 000')) +
  annotate(geom='text', label='India', x=80, y=-5) +
  annotate(geom='text', label='Pakistan', x=65, y=0) +
  annotate(geom='text', label='China', x=150, y=37, hjust=0) +
  annotate(geom='text', label='Bangladesh', x=135, y=23, hjust=0) +
  annotate(geom='text', label='Philippines', x=147, y=17, hjust=0) +
  annotate(geom='text', label='Indonesia', x=90, y=-15, hjust=0) +
  annotate(geom='text', label='Nigeria', x=-20, y=-5, hjust=0) +
  #annotate(geom='text', label='South Africa', x=10, y=-45, hjust=0) +
  annotate(geom='text', label='Democratic Republic of the Congo', x=-30, y=-40, hjust=0) +
  geom_segment(x=80, xend=80, y=-3, yend=20) +
  geom_segment(x=65, xend=65, y=3, yend=28) +
  geom_segment(x=110, xend=145, y=37, yend=37) +
  geom_segment(x=92, xend=132, y=23, yend=23) +
  geom_segment(x=122, xend=145, y=17, yend=17) +
  geom_segment(x=105, xend=115, y=-12, yend=-3) +
  geom_segment(x=-8, xend=8, y=-2, yend=9) +
  #geom_segment(x=25, xend=25, y=-42, yend=-30) 
  #DRC segment below
  geom_segment(x=10, xend=25, y=-35, yend=-3) 


output_ggplot(f1.1.2_plot, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

```


<hr />
<br />

The severity of national TB epidemics, in terms of the number of incident TB cases per `r cm` population in `r yr`, varies widely among countries: from less than ten to more than 500 cases per `r cm` population  (`r lnk("Fig. 1.1.3")`). 

Countries with the highest annual rates were mostly in the African Region. There were `r count(est,year==yr & inc<10)[[2,2]]` countries with a low incidence of TB (<10 cases per `r cm` population per year), mostly in the WHO Region of the Americas and the European Region, plus a few countries in the WHO Eastern Mediterranean and Western Pacific regions.

### `r anch("Fig. 1.1.3")`<span class="red">Fig. 1.1.3 </span>Estimated TB incidence rates, `r yr`

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1.3, fig.alt="Estimated TB incidence rates, 2023"} 


dta <- est[year==yr, .(iso3, inc, country)]
dta$var <- cut(
  dta$inc,
  c(0, 10, 50, 100, 300, 500, Inf),
  c('0-9.9', '10\u201349', '50\u201399', '100\u2013299', '300\u2013499', '\u2265500'),
  right = F,
  ordered_result = T
)

f1.1.3_plot <- whomap(
  X = dta, water.col = 'white',
  legend.title =  "Incidence\nper 100 000\npopulation per year",
  colours = brewer.pal(6, "Greens"),
  legend.pos = c(0.14, 0.34)
)

output_ggplot(f1.1.3_plot, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)


```


<hr />
<br />

In `r yr`, the proportion of people with a new episode of TB (either new or relapse cases) who were living with HIV was highest in countries in the WHO African Region, exceeding 50% in parts of southern Africa (`r lnk("Fig. 1.1.4")`). 

### `r anch("Fig. 1.1.4")`<span class="red">Fig. 1.1.4</span> Estimated HIV prevalence in people with a new episode of TB (new or relapse cases), `r yr`

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1.4, fig.alt="Estimated HIV prevalence in people with new or relapse TB, 2023"} 


dta <- est[year==yr, .(iso3, tbhiv)]
dta$var <- cut(
  dta$tbhiv * 100,
  c(0, 5, 10, 20, 50, Inf),
  c('0\u20134.9', '5\u20139.9', '10\u201319', '20\u201349', '\u226550'),
  right = F,
  ordered_result = T
)

f1.1.4_plot <- whomap(
  X = dta, water.col = 'white',
  colours = brewer.pal(5, "Reds"),
  legend.pos = c(0.14, 0.34),
  legend.title =  "HIV prevalence,\nall ages (%)"
  ) 

output_ggplot(f1.1.4_plot, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

```


<hr /> 
<br />

TB can affect anyone, regardless of age or sex (`r lnk("Fig. 1.1.5")`). The highest burden is in adult men (aged &#8805;15 years), with an estimated  `r round(inc.men/1e6,1)`.0 million cases (95% UI: `r round(inc.men.lo/1e6,1)`–`r round(inc.men.hi/1e6,1)` million) in `r yr`, equivalent to  `r round(100*sum(globsplt$inc[globsplt$sex=="M"&globsplt$child==F])/global$inc.num[global$year==yr],0)`% of the estimated total; by comparison, there were an estimated `r round(inc.women/1e6,1)` million cases (95% UI: `r round(inc.women.lo/1e6,1)`–`r round(inc.women.hi/1e6,1)` million) among adult women (aged &#8805;15 years), equivalent to `r round(100*sum(globsplt$inc[globsplt$sex=="F"&globsplt$child==F])/global$inc.num[global$year==yr],0)`% of the estimated total, and `r round(inc.child/1e6,2)` million (95% UI: `r round(inc.child.lo/1e6,2)`–`r round(inc.child.hi/1e6,2)` million) cases among children (aged 0–14 years), equivalent to `r round(100*sum(globsplt$inc[globsplt$child==T])/global$inc.num[global$year==yr],0)`% of the estimated total. Case notifications of people newly diagnosed with TB in `r yr` were `r round(100*global$c.newinc[global$year==yr]/global$inc.num[global$year==yr],0)`% of the estimated TB incidence overall but this varied by age and sex: `r round(100*sum(globsplt$newrel[globsplt$sex=="M"&globsplt$child==F])/sum(globsplt$inc[globsplt$sex=="M"&globsplt$child==F]),0)`% for adult men, `r round(100*sum(globsplt$newrel[globsplt$sex=="F"&globsplt$child==F])/sum(globsplt$inc[globsplt$sex=="F"&globsplt$child==F]),0)`% for adult women and `r round(100*sum(globsplt$newrel[globsplt$child==T])/sum(globsplt$inc[globsplt$child==T]),0)`% for children aged 0–14 years. Among children aged <5 years, the figure was only `r round(100*sum(globsplt$newrel[globsplt$age=="0-4"])/sum(globsplt$inc[globsplt$age=="0-4"]),0)`%. 


### `r anch("Fig. 1.1.5")`<span class="red">Fig. 1.1.5</span> Global estimates of TB incidence (absolute numbers) disaggregated by age group and sex, compared with case notifications (female in <span style="color:#6363C0">purple</span>; male in <span style="color:#F4A81D">orange</span>), `r yr`


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1.5, fig.alt="Global estimates of TB incidence numbers and case notifications disaggregated by age and sex (female in purple; male in orange), 2023"} 

absspace <- function(x,...) {             #works
     format(abs(x), ..., big.mark=" ",scientific = FALSE, trim = TRUE)
 }
 
 ## ages and colors etc -- From Pete, with a minor edit
 agz3 <- c('0\u20134',
           '5\u201314',
           '15\u201324',
           '25\u201334',
           '35\u201344',
           '45\u201354',
           '55\u201364',
           '65plus') #for labels
 agz4 <- c(rev(rev(agz3)[-1]), "\u226565")
 clz <- c('F' = palette_gtb('female'),
          'M' = palette_gtb('male'))
 ## plot construction
 f1.1.5_plot  <- ggplot(data = globsplt, aes(x = age, y = newrel, fill = sex)) +
   coord_flip() +
   geom_bar(stat = 'identity', aes(y = ifelse(sex == 'M',
                                              newrel,
                                              -newrel))) +
   scale_fill_manual(values = clz) +
   geom_bar(stat = 'identity',
            fill = 'transparent',
            col = 1,
            aes(y = ifelse(sex == 'M',
                           inc,
                           -inc))) +
   theme_gtb(axis_text_size = 12) +
   theme(legend.position = "none") + xlab('Age group (years)') + ylab('Number of TB cases') +
   scale_x_discrete(labels = agz4) +
   scale_y_continuous(labels = absspace) +
   theme(axis.title.y = element_text(margin = margin(r = 15))) 
 
output_ggplot(f1.1.5_plot, globsplt, show_static, pdf_csv_folder, save_csv, save_pdf=FALSE, save_cairo=TRUE)

# Produce the alternative ggplot2 version with estimates as horizontal ribbons with uncertainty intervals 
# (one for each sex) and without showing notifications. This will be used in the main report PDF as Figure 4.
#
# Use the db_estimates_group data table because that also contains disaggregations by age group and sex

FIG.4_data <- db_estimates_group |> 
  dplyr::filter(group_type=="global" & 
                  year == 2023 & 
                  measure == "inc" & 
                  unit == "num" & 
                  sex %in% c("m", "f") & 
                  age_group %in% c("0-4", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65plus")) |> 
  dplyr::select(age_group, sex, best, lo, hi)

FIG.4_data$age_group <- factor(FIG.4_data$age_group,
                               levels=c("0-4", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65plus"),
                               labels=c("0\u20134", "5\u201314", "15\u201324", "25\u201334", "35\u201344", "45\u201354", "55\u201364", "\u226565"))

sex_colours <-  c("f" = gtbreport::palette_gtb("female"), "m" = gtbreport::palette_gtb("male"))
sex_names <- c("Females", "Males") 

FIG.4_plot <- FIG.4_data |> 
  
  ggplot() + 
  
  geom_line(aes(x=age_group, y=best, group=sex, colour=sex), size=1) +
  
  geom_ribbon(aes(x=age_group, ymin=lo, ymax=hi, group=sex, fill = sex), alpha = 0.4) +
  
  scale_colour_manual(values = sex_colours, labels = sex_names) +
  
  scale_fill_manual(values = sex_colours, labels = sex_names) +
  
  # display y-axis scale in millions
  scale_y_continuous(name = "Number of TB cases (millions)", 
                     limits = c(0, 1.4e6),
                     # Use the remainder operator in the labeller function to make sure we don't get weird effects
                     # when plotting small numbers
                     labels = function(i){ifelse((i/1e6) %% 1 == 0, round(i/1e6), round(i/1e6, 1))}) +

  xlab("Age group (years)") +
  
  theme_gtb()  +
  
  theme(legend.position = "none") +
  
  # Get rid of annoying x-axis line 
  theme(axis.line.x = ggplot2::element_blank())

output_ggplot(FIG.4_plot, FIG.4_data, show_static, pdf_csv_folder, save_csv, save_pdf=FALSE, save_cairo=TRUE)


```

<div id="fig_1_1_5_global"></div>

<hr />
<br />

Regional estimates of the distribution of incident TB cases, disaggregated by age and sex and compared with case notifications, are shown in `r lnk("Fig. 1.1.6")`. The largest numbers are typically in the age groups of 25–34, 35–44 or 45–54 years. The WHO Western Pacific Region is an exception, with the largest number of cases in the age group &#8805;65 years; this is mostly influenced by numbers for China, and can be explained by the country’s aging population and declining burden of TB disease (with higher rates of infection and a higher risk of new cases arising from reactivation of previous infections among older people). 

### `r anch("Fig. 1.1.6")`<span class="red">Fig. 1.1.6</span> Regional estimates of TB incidence (absolute numbers) disaggregated by age group and sex, compared with case notifications (female in <span style="color:#6363C0">purple</span>; male in <span style="color:#F4A81D">orange</span>), `r yr`


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=10, fig_1.1.6, fig.alt="Regional estimates of TB incidence numbers and case notifications disaggregated by age and sex, 2023"} 

agz3 <- c('0\u20134',
           '5\u201314',
           '15\u201324',
           '25\u201334',
           '35\u201344',
           '45\u201354',
           '55\u201364',
           '65plus') #for labels
 agz4 <- c(rev(rev(agz3)[-1]), "\u226565")
 clz <- c('F' = palette_gtb('female'),
          'M' = palette_gtb('male'))
 
f1.1.6_plot <- ggplot(data = regsplt, aes(x = age, y = newrel/1000, fill = sex)) +
   coord_flip() +
   geom_bar(stat = 'identity', aes(y = ifelse(sex == 'M',
                                              newrel/1000,
                                              -newrel/1000))) +
   scale_fill_manual(values = clz) +
   geom_bar(stat = 'identity',
            fill = 'transparent',
            col = 1,
            aes(y = ifelse(sex == 'M',
                           inc/1000,
                           -inc/1000))) +
   xlab('Age group (years)') + ylab('') +
   scale_x_discrete(labels = agz4) +
   scale_y_continuous(labels = absspace) +
     facet_wrap(
     ~ name,
     scales = 'free',
     ncol= 3,
     labeller = label_wrap_gen(width = 25)
   ) +
   theme_gtb() +
   theme(legend.pos = 'none')

output_ggplot(f1.1.6_plot, regsplt, show_static, pdf_csv_folder, save_csv, save_pdf=FALSE, save_cairo=TRUE)


```

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_6_AFR"></div>
<div id="fig_1_1_6_EUR"></div>
</div>
	
<div class="col-md-4">
<div id="fig_1_1_6_AMR"></div>
<div id="fig_1_1_6_EMR"></div>
</div>
	
<div class="col-md-4">
<div id="fig_1_1_6_SEAR"></div>
<div id="fig_1_1_6_WPR"></div>
</div>
</div>


<hr />
<br />


Regional trends in TB incidence rates vary, and progress in relation to the `r inc.mort.milestone.yr` milestone of the End TB Strategy is mixed (`r lnk("Fig. 1.1.7")`).


### `r anch("Fig. 1.1.7")`<span class="red">Fig. 1.1.7</span> Trends in estimated TB incidence rates by WHO region, 2010&#8211;`r yr`
<div class="subhead">The overall TB incidence rate is shown in <span style="color:#91A93E">green</span> and the incidence rate among people living with HIV is shown in <span style="color:#ED1D24">red</span>. The black solid lines show case notifications of people newly diagnosed with TB, for comparison with estimates of the overall incidence rate. Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1-inc.milestone.vs.2015)*100`% reduction in the TB incidence rate between 2015 and `r inc.mort.milestone.yr`.</div> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=8,  fig_1.1.7, fig.alt="Trends in estimated TB incidence rates by WHO region, 2000-2023"} 

f1.1.7_plot <- qplot(year,
      inc,
      data = subset(regional,year>=2010),
      geom = 'line',
      colour = I('blue')) +
  geom_line(aes(year, inc.h), colour = I('lightskyblue2')) +
  geom_ribbon(aes(year, ymin = inc.h.lo, ymax = inc.h.hi),
              fill = I('lightskyblue2'),
              alpha = 0.4) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, newinc)) +
  geom_hline(aes(yintercept = inc.milestone), linetype = I(2)) +
  facet_wrap(~ region, scale = 'free_y') +
  xlab('') + ylab('Incidence rate per 100 000 population per year') +
  expand_limits(y=1) +
  theme_bw(base_size = 12) +
  #  scale_y_continuous(trans='log10', breaks=c(1,2,5,10,20,30,50,100,200,300)) +
 # scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme_gtb() +
  theme(legend.position = 'none') 

output_ggplot(f1.1.7_plot, subset(regional,year>=2010), show_static, pdf_csv_folder, save_csv, save_pdf)


```


<div class="row">
<div class="col-md-4">
<div id="fig_1_1_7_AFR"></div>
<div id="fig_1_1_7_EUR"></div>
</div>
	
<div class="col-md-4">
<div id="fig_1_1_7_AMR"></div>
<div id="fig_1_1_7_EMR"></div>
</div>
	
<div class="col-md-4">
<div id="fig_1_1_7_SEAR"></div>
<div id="fig_1_1_7_WPR"></div>
</div>
</div>

<hr />
<br />


In 2023, the TB incidence rate increased for the third consecutive year in the Region of the Americas and the Western Pacific Region, reflecting the estimated impact of shortfalls in TB case detection in 2020 and incomplete recovery in 2021 (in both regions) as well as 2022 (in the Western Pacific). An upward trend in 2020–2022 was reversed in 2023 in two regions: the Eastern Mediterranean Region and the South-East Asian Region. In the European Region, the TB incidence rate started to decrease in 2022. In the African Region, the decline that has been sustained for many years was maintained in 2023, consistent with evidence that disruptions related to the COVID-19 pandemic had little impact on the number of people diagnosed and officially notified with TB (<span class="red">Section 2.1</span>).

The biggest estimated reduction in the TB incidence rate compared with the baseline year of 2015 has been in the European Region, with a net reduction of `r round(100-100*regional$inc[regional$year==yr & regional$g.whoregion=="EUR"]/regional$inc[regional$year==2015 & regional$g.whoregion=="EUR"],0)`% by `r yr`. A similar reduction has been achieved in the African Region, with a `r round(100-100*regional$inc[regional$year==yr & regional$g.whoregion=="AFR"]/regional$inc[regional$year==2015 & regional$g.whoregion=="AFR"],0)`% decline by `r yr`. The net decline by `r yr` compared with 2015 was relatively small in three of the remaining WHO regions: the Eastern Mediterranean (`r round(100-100*regional$inc[regional$year==yr & regional$g.whoregion=="EMR"]/regional$inc[regional$year==2015 & regional$g.whoregion=="EMR"],1)`%), South-East Asia (`r round(100-100*regional$inc[regional$year==yr & regional$g.whoregion=="SEA"]/regional$inc[regional$year==2015 & regional$g.whoregion=="SEA"],1)`%) and the Western Pacific (`r round(100-100*regional$inc[regional$year==yr & regional$g.whoregion=="WPR"]/regional$inc[regional$year==2015 & regional$g.whoregion=="WPR"],1)`%). The most concerning trend is in the Region of the Americas, where there was a net increase of `r round(100*regional$inc[regional$year==yr & regional$g.whoregion=="AMR"]/regional$inc[regional$year==2015 & regional$g.whoregion=="AMR"]-100,0)`% between 2015 and `r yr`.  



Trends in TB incidence rates by income group are shown in (`r lnk("Fig. 1.1.8")`).


### `r anch("Fig. 1.1.8")`<span class="red">Fig. 1.1.8</span> Trends in estimated TB incidence rates by income group, 2010&#8211;`r yr`

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=8,  fig_1.1.8, fig.alt="Trends in estimated TB incidence rates by WHO region, 2000-2023"} 

f1.1.8_plot <- qplot(year,
      inc,
      data = subset(income,year>=2010),
      geom = 'line',
      colour = I('blue')) +
  geom_line(aes(year, inc.h), colour = I('lightskyblue2')) +
  geom_ribbon(aes(year, ymin = inc.h.lo, ymax = inc.h.hi),
              fill = I('lightskyblue2'),
              alpha = 0.4) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, newinc)) +
  facet_wrap(~ group, scale = 'free_y') +
  xlab('') + ylab('Incidence rate per 100 000 population per year') +
  expand_limits(y=1) +
  theme_bw(base_size = 12) +
  #  scale_y_continuous(trans='log10', breaks=c(1,2,5,10,20,30,50,100,200,300)) +
 # scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme_gtb() +
  theme(legend.position = 'none') 

output_ggplot(f1.1.8_plot, subset(income,year>=2010), show_static, pdf_csv_folder, save_csv, save_pdf)


```



<div class="row">
<div class="col-md-6">
<div id="fig_1_1_8_LIC"></div>
<div id="fig_1_1_8_LMC"></div>
</div>
	
<div class="col-md-6">
<div id="fig_1_1_8_UMC"></div>
<div id="fig_1_1_8_HIC"></div>
</div>
</div>

<hr />
<br />




Progress in reducing the TB incidence rate at country level is highly variable (`r lnk("Fig. 1.1.9")`). Many countries, especially in the European and African regions, have achieved estimated reductions of more than 20% between 2015 and `r yr`, thus surpassing the first (2020) milestone of the End TB Strategy. More than 10 countries are estimated to have achieved reductions of at least 50% between 2015 and `r yr`, including one high TB burden country (South Africa). At the other extreme, there are countries where the TB incidence rate in `r yr` was more than 5% higher than in 2015. These countries are most noticeably in the Region of the Americas but also include four high TB burden countries in Asia: Indonesia, Mongolia, Myanmar and the Philippines. 





### `r anch("Fig. 1.1.9")`<span class="red">Fig. 1.1.9</span> Change (%) in estimated TB incidence (new cases per `r cm` population), `r yr` compared with 2015

<div class="subhead">The first milestone of the End TB Strategy was a 20% reduction by 2020, compared with 2015; the second milestone is a 50% reduction by 2025, compared with 2015. The last two categories (decrease 20-49%, and decrease &#8805;50%) distinguish the countries that have made the most progress towards the second milestone of the End TB Strategy.</div>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1.9, fig.alt="Change in TB incidence between 2015 and 2023"} 

dta <- subset(est, year==yr, select=c("iso3","country","year","inc"))
dta_2015 <- subset(est, year==2015, select=c("iso3","inc"))
setnames(dta_2015, "inc", "inc.2015")
dta <- merge(dta, dta_2015, by="iso3")
rm(dta_2015)

dta$rel.red.inc=(1-(dta$inc/dta$inc.2015))*100
dta$rel.red.inc.cat=ifelse(dta$rel.red.inc>0,"Reduction","Increase")

dta$rel.red.inc.fordisplay=ifelse(dta$rel.red.inc<0,
                                      -round(dta$rel.red.inc,1),
                                      round(dta$rel.red.inc,1))

dta$var <- cut(
  dta$rel.red.inc,
  c(-Inf,-5,5, 10, 20, 50,Inf),
  c("Increase (>5%)",'Stable (-5% to +5%)', 'Decrease (6\u20139%)', 'Decrease (10\u201319%)', 'Decrease (20\u201349%)','Decrease (\u226550%)'),
  right = F,
  ordered_result = T
)

f1.1.9_plot <- whomap(
  X = dta, water.col = 'white',
  legend.title = paste0("Change (", yr, " vs 2015, %)"),
  legend.pos = c(0.14, 0.34),
  colours = c("lightyellow2","lightyellow3",brewer.pal(4, "Purples"))
)

output_ggplot(f1.1.9_plot, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

```


<hr />
<br />

Trends in the TB incidence rate in the 30 high TB burden countries are also highly variable (`r lnk("Fig. 1.1.10")`). In `r yr`, the best estimate of the TB incidence rate suggested that the 2025 milestone of the End TB Strategy had been reached in South Africa and was close to being reached in Kenya, the United Republic of Tanzania and Zambia.


### `r anch("Fig. 1.1.10")`<span class="red">Fig. 1.1.10</span> Trends in estimated TB incidence rates in the 30 high TB burden countries compared with notifications of new and relapse cases, 2010&#8211;`r yr`
<div class="subhead">TB incidence rates are shown in <span style="color:#91A93E">green</span>. The black solid lines show notifications of new and relapse cases for comparison with estimates of the total incidence rate. Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1-inc.milestone.vs.2015)*100`% reduction between 2015 and `r inc.mort.milestone.yr`.</div> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=12, fig_1.1.10, fig.alt="Trends in estimated TB incidence rates in the 30 high TB burden countries compared with notifications of new and relapse cases, 2010−2023"} 


f1.1.10_plot <- qplot(year,
      inc,
      data = subset(hest,year>=2010),
      geom = 'line',
      colour = I('blue')) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, newinc)) +
  geom_hline(aes(yintercept = inc.milestone), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
  facet_wrap(~country, nrow=6, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
#  expand_limits(y = 1) +
  xlab('') + ylab('Incidence rate per 100 000 population per year') +
 # scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme_gtb()


output_ggplot(f1.1.10_plot, subset(hest,year>=2010), show_static, pdf_csv_folder, save_csv, save_pdf)

```

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_AGO"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_BGD"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_BRA"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_CAF"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_CHN"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_COG"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_PRK"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_COD"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_ETH"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_GAB"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_IND"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_IDN"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_KEN"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_LSO"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_LBR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_MNG"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_MOZ"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_MMR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_NAM"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_NGA"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_PAK"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_PNG"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_PHL"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_SLE"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_ZAF"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_THA"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_UGA"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_9_TZA"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_VNM"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_9_ZMB"></div>
</div>
</div>


<hr />
<br />

In 2021, 2022 and 2023, the estimated TB incidence rate increased in two of the three global TB watchlist countries, while the Russian Federation is close to reaching the 2025 milestone of the End TB Strategy (`r lnk("Fig. 1.1.11")`). 


### `r anch("Fig. 1.1.11")`<span class="red">Fig. 1.1.11</span> Trends in estimated TB incidence rates in the three global TB watchlist countries compared with notifications of new and relapse cases, 2010&#8211;`r yr`
<div class="subhead">TB incidence rates are shown in <span style="color:#91A93E">green</span>. The black solid lines show notifications of new and relapse cases for comparison with estimates of the total incidence rate. Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1-inc.milestone.vs.2015)*100`% reduction between 2015 and `r inc.mort.milestone.yr`.</div> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=2.7, fig.width=8, fig.align='left', fig_1.1.11, fig.alt="Trends in estimated TB incidence rates in the 3 WHO watchlist countries compared with notifications of new and relapse cases, 2010−2023"} 

dta <- est[iso3 %in% c('KHM','RUS','ZWE')]

f1.1.11_plot <- qplot(year,
      inc,
      data = subset(dta,year>=2010),
      geom = 'line',
      colour = I('blue')) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, newinc)) +
  geom_hline(aes(yintercept = inc.milestone), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
  facet_wrap(~country, nrow=1, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
#  expand_limits(y = 1) +
  xlab('') + ylab('Incidence rate per 100 000\npopulation per year') +
 # scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme_gtb()

output_ggplot(f1.1.11_plot, subset(dta,year>=2010), show_static, pdf_csv_folder, save_csv, save_pdf)


```

<div class="row">
<div class="col-md-4">
<div id="fig_1_1_10_KHM"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_10_RUS"></div>
</div>
<div class="col-md-4">
<div id="fig_1_1_10_ZWE"></div>
</div>
</div>

<hr />
<br />



`r anch("Box 1.1.1")`


<div class="textbox">
## Box 1.1.1 

## Methods used by WHO to estimate TB incidence 

The main methods used by WHO to estimate TB incidence at country level in the period 2010–`r yr` are shown in `r lnk("Fig. 1.1.12")`. These methods adhere to global guidelines for accurate and transparent reporting of health estimates (`r ref_lnk("6")`) and are described in detail in a <span class="red">technical appendix</span>.

Methods used by WHO to estimate TB incidence for 2000–2019 included: 

* results from national TB prevalence surveys  combined with estimates of the duration of disease, used for 28 countries with 39% of the global number of incident TB cases in 2019;

* results from a national TB prevalence survey combined with a country-specific dynamic model that was calibrated to survey data as well as other data (e.g. notifications, drug sales) and WHO TB mortality estimates. This was used for India, which had 27% of the global number of incident TB cases in 2019;

* notifications adjusted by a standard factor to account for underreporting, overdiagnosis and underdiagnosis, used for 137 countries (including most high-income countries and selected middle-income countries and areas) with about 6.0% of the global number of incident TB cases in 2019;

* results from national inventory studies that measured the level of underreporting of detected TB cases, used for 10 countries with about 17% of the global number of incident TB cases in 2019; and

* case notification data combined with expert opinion about case-detection gaps, used for 39 countries with 11% of the global number of incident TB cases in 2019.


For 47 countries with the biggest absolute reductions in TB notifications during the COVID-19 pandemic that were a clear departure from historic trends, estimates of TB incidence in 2020–2023 were based on country or region-specific dynamic models (n=24 and n=23, respectively). These methods are described in detail elsewhere (`r ref_lnk("1, 2, 3")`). For all other countries, the same methods as those used for the period 2010–2019 continued to be applied.


The <span style="color:#f4a81d">orange</span> dots on the map show the 24 countries for which a country-specific dynamic model has been used and the  <span style="color:#754d06">brown</span> dots show the 23 countries for which a region-specific dynamic model has been used.  These two groups of countries accounted for `r 62`% and `r 0.7`% of the estimated total global number of incident cases in `r yr`, respectively.

<hr />
<br />

### `r anch("Fig. 1.1.12")`<span class="red">Fig. 1.1.12</span> Main methods used to estimate TB incidence, 2010&#8211;`r yr`

<!-- <div class="subhead"> The <span style="color:#f4a81d">orange</span> dots are showing countries for which a country-specific dynamic model has been used to estimate TB incidence and the <span style="color:#754d06">green</span> dots are showing countries for which a region-specific dynamic model has been used to estimate TB incidence. </div> -->


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1.12, fig.alt="Main methods used to estimate TB incidence, 2010–2023"} 

# dta <-
#   est[year == 2019, .(iso3,
#                         source.inc)]
# dta$source.inc[dta$iso3=="IND"] = "Prevalence survey and country model"
# dta$var <- dta$source.inc
# 
# dta$var <- plyr::revalue(dta$var, c("Country model"="Country-specific dynamic model", 
#                               "Regional model"="Region-specific dynamic model", 
#                               "Case notifications, standard adjustment"="Notifications and standard adjustment",
#                               "Case notifications, Standard adjustment"="Notifications and standard adjustment",
#                               "Case notifications,\nExpert opinion"="Notifications and expert opinion",
#                               "Case notifications, expert opinion"="Notifications and expert opinion"))
# 
# 
# method1=dta
# 
# f1.1.12_plot <- whomap(
#   X = method1, water.col = 'white',
#   legend.title = 'Main method',
#   #colours = brewer.pal(4, "Paired"),
#   colours= c("#fbb4b9","#bdd7e7","#2c7fb8","#c51b8a","#88be3b"),
#   legend.pos = c(0.11,0.48)
#   )
# 
# 
# dta <-
#   est[year == yr, .(iso3,source.inc)]
# dta$var <- dta$source.inc
# 
# dta$var <- plyr::revalue(dta$var, c("Country model"="Country-specific dynamic model", 
#                               "Regional model"="Region-specific dynamic model", 
#                                                             "Case notifications, standard adjustment"="Notifications and standard adjustment",
#                               "Case notifications, Standard adjustment"="Notifications and standard adjustment",
#                               "Case notifications,\nExpert opinion"="Notifications and expert opinion",
#                               "Case notifications, expert opinion"="Notifications and expert opinion"))
# 
# method2=dta

# 
# f1.1.13_plot <- whomap(
#   X = method2, water.col = 'white',
#   legend.title = 'Main method',
#   #colours = brewer.pal(4, "Dark2"),
#   colours= c("#bdd7e7","#2c7fb8", "#88be3b", "#fbb4b9", "#ffffcc", "#c51b8a", "#308047"),
#   legend.pos = c(0.11, 0.35)
# )
# 

cm.lst=unique(est$iso3[est$meth.inc23=="Country-specific dynamic model"])
rm.lst=unique(est$iso3[est$meth.inc23=="Region-specific dynamic model"])
rm.lst=c(rm.lst,"ARG","CUB","ECU","GTM","PRY","GRD")

method1=est[year==2019,.(iso3,var=meth.inc19)]

method1["FRA",var:="Inventory study"]
method1["KOR",var:="Inventory study"]

f1.1.12_plot <- whomap(
  X = method1, water.col = 'white',
  legend.title = 'Main method',
  #colours = brewer.pal(4, "Paired"),
  colours= c("#fbb4b9","#bdd7e7","#2c7fb8","#c51b8a","#88be3b"),
  legend.pos = c(0.11,0.48)
  )


f1.1.12_plot  <- f1.1.12_plot +
  add_marker(iso3=cm.lst, col="#f4a81d", size=2, shape=16) + 
  add_marker(iso3=rm.lst, col="#754d06", size=2, shape=16) + 
  guides(shape=FALSE) +
  annotate("text", label = "Country-specific dynamic model", x = -120, y = -48,size = I(2.5)) +
  annotate("point",  x = -166, y = -48,color="#f4a81d",size = I(2)) +
  annotate("text", label = "Region-specific dynamic model", x = -120, y = -53,size = I(2.5)) +
  annotate("point",  x = -166, y = -53,color="#754d06",size = I(2))


output_ggplot(f1.1.12_plot, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

```

</div>


Country-specific details are available in the [Global tuberculosis report app](https://www.who.int/teams/global-tuberculosis-programme/data#app) and [country profiles](https://worldhealthorg.shinyapps.io/tb_profiles/).


Note: All of the percentage reductions or increases referred to in the text of this webpage were calculated based on unrounded numbers, as opposed to the rounded numbers that appear in the interactive graphics.

`r anch("refs")`

<hr style="border:1px solid gray20">

**References**

1. Global tuberculosis report 2021. Geneva: World Health Organization; 2021 (https://iris.who.int/handle/10665/346387).

2. Global tuberculosis report 2022. Geneva: World Health Organization; 2022 (https://iris.who.int/handle/10665/363752).

3. Global tuberculosis report 2023. Geneva: World Health Organization; 2023 (https://iris.who.int/handle/10665/373828).

4. Floyd K, Glaziou P, Houben R, Sumner T, White RG, Raviglione M. Global tuberculosis targets and milestones set for 2016-2035: definition and rationale. Int J Tuberc Lung Dis. 2018;22(7):723-30 (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6005124/). 

5. World Health Organization. Resolution WHA67.1. Global strategy and targets for tuberculosis prevention, care and control after 2015. Geneva: World Health Organization; 2014 (http://apps.who.int/gb/ebwha/pdf_files/WHA67/A67_R1-en.pdf). 

6. Guidelines for Accurate and Transparent Health Estimates Reporting (GATHER) [website] (http://gather-statement.org/). 
    
  



```{r js_functions}
# Insert javascript file containing common Kendo number formatting functions ----
cat(writeLines(readLines(here("report/resources/gtbr_js.htm"))))
```



```{r, json_2_1} 
# renaming variable for Kendo UI
names(global) <- gsub('[.]', '_', names(global))

names(regional) <- gsub('[.]', '_', names(regional))

names(hest) <- gsub('[.]', '_', names(hest))

names(income) <- gsub('[.]', '_', names(income))


dta <- est[iso3 %in% c('KHM','RUS','ZWE')]
names(dta) <- gsub('[.]', '_', names(dta))

```

<script type="text/javascript">

/* JSON data objects for the figures */


var fig_1_1_1a_data = `r subset(global,year>=2010) %>% dplyr::select(year,inc_num,inc_lo_num,inc_hi_num,c_newinc,inc_h_num,inc_h_hi_num,inc_h_lo_num) %>% toJSON("rows")`; 
var fig_1_1_1b_data = `r subset(global,year>=2010) %>% dplyr::select(year,inc,inc_lo,inc_hi,newinc,inc_h,inc_h_hi,inc_h_lo,inc_milestone) %>% toJSON("rows")`; 

var fig_1_1_5_data = `r globsplt %>% pivot_wider(names_from = sex, values_from = inc:pop) %>% mutate(inc_F = inc_F*(-1), newrel_F = newrel_F*(-1)) %>% mutate(diff_M = inc_M-newrel_M, diff_F=inc_F-newrel_F) %>% arrange(rev(age))%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_1_6_data = `r regsplt %>% select(!pop) %>% pivot_wider(names_from = sex, values_from = inc:newrel) %>% mutate(inc_F = inc_F*(-1), newrel_F = newrel_F*(-1)) %>% mutate(diff_M = inc_M-newrel_M, diff_F=inc_F-newrel_F) %>% arrange(rev(age)) %>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 


var fig_1_1_7_data = `r subset(regional,year>=2010) %>% dplyr::select(region,year,inc,inc_lo,inc_hi,newinc,inc_h,inc_h_hi,inc_h_lo,inc_milestone) %>% toJSON("rows")`; 

var fig_1_1_8_data = `r subset(income,year>=2010) %>% dplyr::select(group,year,inc,inc_lo,inc_hi,newinc,inc_h,inc_h_hi,inc_h_lo) %>% toJSON("rows")`; 

var fig_1_1_9_data = `r subset(hest,year>=2010) %>% dplyr::select(country,iso3,year,inc,inc_lo,inc_hi,newinc,inc_h,inc_h_hi,inc_h_lo,inc_milestone) %>% toJSON("rows")`; 

var fig_1_1_10_data = `r subset(dta,year>=2010) %>% dplyr::select(country,iso3,year,inc,inc_lo,inc_hi,newinc,inc_h,inc_h_hi,inc_h_lo,inc_milestone) %>% toJSON("rows")`; 

</script>


```{js, echo=FALSE}

function tb_format_thou_2(n) { 
  //return thousands formatted to to significant figures
  nt = n/1000; 
  return num_spacer(Number(nt.toPrecision(2))*1000)
}

/* Functions to create the figures */

function createFig_1_1_1a(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_1_1_1a_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "All TB cases",
				field: "inc_num",
				color: "#91A93E",
				tooltip: {
					visible: true,
					template: "All TB cases (#= category #): #= tb_format_thou(value) # 000"
				}
			},{
				type: "rangeArea",
				fromField: "inc_lo_num",
				toField: "inc_hi_num",
				color: "#91A93E",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= tb_format_thou(value.from) # 000–#= tb_format_thou(value.to) # 000"
				}
			},{
				type: "line",
        name: "Notifications of new and relapse cases",
				field: "c_newinc",
				color: "black",
				tooltip: {
					visible: true,
					format: "{0}",
					template: "Notifications of new and relapse cases (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "line",
        name: "TB cases in people living with HIV",
				field: "inc_h_num",
				color: "#ED1D24",
				tooltip: {
					visible: true,
					template: "TB cases in people living with HIV (#= category #): #= tb_format_thou(value) # 000"
				}
			},{
				type: "rangeArea",
				fromField: "inc_h_lo_num",
				toField: "inc_h_hi_num",
				color: "#ED1D24",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= tb_format_thou(value.from) # 000–#= tb_format_thou(value.to) # 000"
				}
			},],      
			valueAxis: {

        labels: {
					template: "#= axis_spacer(value/1e6) #"
				},
				title: {
					text: "Millions per year",
					visible: true
				},
        min: 0,

				line: {
					visible: false
				}
			}
			,
			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}



function createFig_1_1_1b(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_1_1_1b_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "All TB cases",
				field: "inc",
				color: "#91A93E",
				tooltip: {
					visible: true,
					template: "All TB cases (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "inc_lo",
				toField: "inc_hi",
				color: "#91A93E",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= num_spacer(value.from) #–#= num_spacer(value.to) #"
				}
			},{
				type: "line",
        name: "Notifications of new and relapse cases",
				field: "newinc",
				color: "black",
				tooltip: {
					visible: true,
					format: "{0}",
					template: "Notifications of new and relapse cases (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "line",
        name: "TB cases in people living with HIV",
				field: "inc_h",
				color: "#ED1D24",
				tooltip: {
					visible: true,
					template: "Incidence rate of TB in people living with HIV (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "inc_h_lo",
				toField: "inc_h_hi",
				color: "#ED1D24",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= num_spacer(value.from) #–#= num_spacer(value.to) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "inc_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false,
			}
      },],      
      
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Rate per 100 000 population per year",
					visible: true
				},
        min: 0,
        max: 250,
				line: {
					visible: false
				}
			},      
      render: function (e) {
            var draw = kendo.drawing;
            var padding = 100;
            var padding2 = 300;
            var element = this.element;
            var rect = new kendo.geometry.Rect([padding, padding2], [element.width() - 2 * padding, element.height() - 2 * padding2]);

            var text = new draw.Text("2025 milestone", [0, 0], { font: "bold 9px Verdana,Arial,sans-serif" });
            draw.align([text], rect, "start");
            draw.vAlign([text], rect, "start");
            e.sender.surface.draw(text);
          },
			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}


function createFig_1_1_5(fig_ID, data, min) {
  
		$(fig_ID).kendoChart({		
      dataSource: fig_1_1_5_data,			
      chartArea: {
				height: 500
			},	  
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true,
        gap: 0.2
			},
			series: [{
				field: "newrel_M",
				color: "#F4A81D",
        tooltip: {
				visible: true,
				template: "Number of cases notified in males aged #= category # years: #= num_spacer(value)#"
			}
			},{
				field: "diff_M",
				color: "#EEEEEE",
        tooltip: {
				visible: true,
				template: "Estimated TB incidence number in males aged #= category # years: #= tb_format_thou_2(dataItem.inc_M)#"
			}
			},{
				field: "newrel_F",
				color: "#6363C0",
        tooltip: {
				visible: true,
				template: "Number of cases notified in females aged #= category # years: #= num_spacer(value*(-1))#"
			}
			},{
				field: "diff_F",
				color: "#EEEEEE",
        tooltip: {
				visible: true,
				template: "Estimated TB incidence number in females aged #= category # years: #= tb_format_thou_2(dataItem.inc_F*(-1))#"
			}
			},
              ],
			valueAxis: {
				labels: {
				template:  "#= kendo.toString(Math.abs(value),'n0').replace(/,/g, ' ') #",
				},
        title: {
					text: "Number of TB cases"
				},
				line: {
					visible: false
				},
        axisCrossingValue: min,
        min: min
			},
			categoryAxis: {
				field: "age",
				labels: {
					rotation: "auto"
				},
        title: {
					text: "Age group (years)"
				},
				majorGridLines: {
					visible: false
				}			}
		});
}

function createFig_1_1_6(fig_ID, data, filter, min) {
  
  	// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.name == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
      chartArea: {
				height: 250
			},		     
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true,
        gap: 0.2
			},
			series: [{
				field: "newrel_M",
				color: "#F4A81D",
        tooltip: {
				visible: true,
				template: "Number of cases notified in males aged #= category # years: #= num_spacer(value)#"
			}
			},{
				field: "diff_M",
				color: "#EEEEEE",
        tooltip: {
				visible: true,
				template: "Estimated TB incidence number in males aged #= category # years: #= tb_format_thou_2(dataItem.inc_M)#"
			}
			},{
				field: "newrel_F",
				color: "#6363C0",
        tooltip: {
				visible: true,
				template: "Number of cases notified in females aged #= category # years: #= num_spacer(value*(-1))#"
			}
			},{
				field: "diff_F",
				color: "#EEEEEE",
        tooltip: {
				visible: true,
				template: "Estimated TB incidence number in females aged #= category # years: #= tb_format_thou_2(dataItem.inc_F*(-1))#"
			}
			},
              ],
			valueAxis: {
				labels: {
				template:  "#= kendo.toString(Math.abs(value),'n0').replace(/,/g, ' ') #",
				step:2
				},
        title: {
					text: "Number of TB cases"
				},
				line: {
					visible: false
				},
        axisCrossingValue: min,
        min: min
			},
			categoryAxis: {
				field: "age",
				labels: {
					rotation: "auto"
				},
        title: {
					text: "Age group (years)"
				},
				majorGridLines: {
					visible: false
				}			}
		});
}  


function createFig_1_1_7(fig_ID, data, filter, min, unit) {
 
  	// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.region == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: 400
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "All TB cases",
				field: "inc",
				color: "#91A93E",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "TB incidence rate (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "inc_lo",
				toField: "inc_hi",
				color: "#91A93E",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= num_spacer(value.from) #–#= num_spacer(value.to) #"
				}
			},{
				type: "line",
        name: "Notifications of new and relapse cases",
				field: "newinc",
				color: "black",     
        markers: { size: 4},
				tooltip: {
					visible: true,
					format: "{0}",
					template: "TB case notification rate (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "line",
        name: "TB cases in people living with HIV",
				field: "inc_h",
				color: "#ED1D24", 
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "Incidence rate of TB in people living with HIV (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "inc_h_lo",
				toField: "inc_h_hi",
				color: "#ED1D24",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= num_spacer(value.from) #–#= num_spacer(value.to) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "inc_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
      
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Rate per 100 000 population\nper year",
					visible: true
				},
        min: 0,
        axisCrossingValue: 0.1,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: 0,
          step: 2
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}





function createFig_1_1_9_10(fig_ID, data, filter, min, max, unit) {
 
  	// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.country == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: 250
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
				field: "inc",
				color: "#91A93E",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "TB incidence rate (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "inc_lo",
				toField: "inc_hi",
				color: "#91A93E",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= num_spacer(value.from) #–#= num_spacer(value.to) #"
				}
			},{
				type: "line",
				field: "newinc",
				color: "black",     
        markers: { size: 4},
				tooltip: {
					visible: true,
					format: "{0}",
					template: "TB case notification rate (#= category #): #= num_spacer(value) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "inc_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
      
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Rate per 100 000 population\nper year",
					visible: true
				},
       min: 0,
        axisCrossingValue: 0.1,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: 0,
          step: 2
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}



function createFig_1_1_8(fig_ID, data, filter, min, unit) {
 
  	// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.group == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: 400
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "All TB cases",
				field: "inc",
				color: "#91A93E",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "TB incidence rate (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "inc_lo",
				toField: "inc_hi",
				color: "#91A93E",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= num_spacer(value.from) #–#= num_spacer(value.to) #"
				}
			},{
				type: "line",
        name: "Notifications of new and relapse cases",
				field: "newinc",
				color: "black",     
        markers: { size: 4},
				tooltip: {
					visible: true,
					format: "{0}",
					template: "TB case notification rate (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "line",
        name: "TB cases in people living with HIV",
				field: "inc_h",
				color: "#ED1D24", 
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "Incidence rate of TB in people living with HIV (#= category #): #= num_spacer(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "inc_h_lo",
				toField: "inc_h_hi",
				color: "#ED1D24",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= num_spacer(value.from) #–#= num_spacer(value.to) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "inc_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
      
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Rate per 100 000 population\nper year",
					visible: true
				},
        min: 0,
        axisCrossingValue: 0.1,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: 0,
          step: 2
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}





```

```{js, echo=FALSE}

/* Create the figures after the document has been loaded */

  $(document).ready(function () {
  
  createFig_1_1_1a("#fig_1_1_1a");
  createFig_1_1_1b("#fig_1_1_1b");
  
  createFig_1_1_5("#fig_1_1_5_global",fig_1_1_5_data,-1e6);
  
  createFig_1_1_6("#fig_1_1_6_EUR",fig_1_1_6_data,"European Region",-20000);
  createFig_1_1_6("#fig_1_1_6_AFR",fig_1_1_6_data,"African Region",-200000);
  createFig_1_1_6("#fig_1_1_6_EMR",fig_1_1_6_data,"Eastern Mediterranean Region",-80000);
  createFig_1_1_6("#fig_1_1_6_AMR",fig_1_1_6_data,"Region of the Americas",-30000);
  createFig_1_1_6("#fig_1_1_6_WPR",fig_1_1_6_data,"Western Pacific Region",-150000);
  createFig_1_1_6("#fig_1_1_6_SEAR",fig_1_1_6_data,"South-East Asia Region",-500000);
  
  createFig_1_1_7("#fig_1_1_7_EUR",fig_1_1_7_data,"European Region",1,10);
  createFig_1_1_7("#fig_1_1_7_AFR",fig_1_1_7_data,"African Region",1,10);
  createFig_1_1_7("#fig_1_1_7_EMR",fig_1_1_7_data,"Eastern Mediterranean Region",0.1,10);
  createFig_1_1_7("#fig_1_1_7_AMR",fig_1_1_7_data,"Region of the Americas",1,4);
  createFig_1_1_7("#fig_1_1_7_WPR",fig_1_1_7_data,"Western Pacific Region",1,10);
  createFig_1_1_7("#fig_1_1_7_SEAR",fig_1_1_7_data,"South-East Asia Region",1,10);
 
  createFig_1_1_8("#fig_1_1_8_HIC",fig_1_1_8_data,"High-income countries",1,10);
  createFig_1_1_8("#fig_1_1_8_LIC",fig_1_1_8_data,"Low-income countries",1,10);
  createFig_1_1_8("#fig_1_1_8_LMC",fig_1_1_8_data,"Lower-middle-income countries",0.1,10);
  createFig_1_1_8("#fig_1_1_8_UMC",fig_1_1_8_data,"Upper-middle-income countries",1,4);
  
  createFig_1_1_9_10("#fig_1_1_9_AGO",fig_1_1_9_data,"Angola",80, 600, 2);
  createFig_1_1_9_10("#fig_1_1_9_COG",fig_1_1_9_data,"Congo",100,1000,2);
  createFig_1_1_9_10("#fig_1_1_9_IND",fig_1_1_9_data,"India",80,1000,2);
  createFig_1_1_9_10("#fig_1_1_9_MNG",fig_1_1_9_data,"Mongolia",70,1000,2);
  createFig_1_1_9_10("#fig_1_1_9_PAK",fig_1_1_9_data,"Pakistan",5,600,4);
  createFig_1_1_9_10("#fig_1_1_9_THA",fig_1_1_9_data,"Thailand",50,600,2);
  
  createFig_1_1_9_10("#fig_1_1_9_BGD",fig_1_1_9_data,"Bangladesh",50,600,2);
  createFig_1_1_9_10("#fig_1_1_9_PRK",fig_1_1_9_data,"Democratic People's Republic of Korea",100,700,2);
  createFig_1_1_9_10("#fig_1_1_9_IDN",fig_1_1_9_data,"Indonesia",20,600,2);
  createFig_1_1_9_10("#fig_1_1_9_MOZ",fig_1_1_9_data,"Mozambique",100,600,2);
  createFig_1_1_9_10("#fig_1_1_9_PNG",fig_1_1_9_data,"Papua New Guinea",100,700,2);
  createFig_1_1_9_10("#fig_1_1_9_UGA",fig_1_1_9_data,"Uganda",100,600,2);
  
  createFig_1_1_9_10("#fig_1_1_9_BRA",fig_1_1_9_data,"Brazil",20,100,2);
  createFig_1_1_9_10("#fig_1_1_9_COD",fig_1_1_9_data,"Democratic Republic of the Congo",100,600,2);
  createFig_1_1_9_10("#fig_1_1_9_KEN",fig_1_1_9_data,"Kenya",100,1500,2);
  createFig_1_1_9_10("#fig_1_1_9_MMR",fig_1_1_9_data,"Myanmar",50,1500,2);
  createFig_1_1_9_10("#fig_1_1_9_PHL",fig_1_1_9_data,"Philippines",100,1000,2);
  createFig_1_1_9_10("#fig_1_1_9_TZA",fig_1_1_9_data,"United Republic of Tanzania",50,1500,2);
  
  createFig_1_1_9_10("#fig_1_1_9_CAF",fig_1_1_9_data,"Central African Republic",50,1000,2);
  createFig_1_1_9_10("#fig_1_1_9_ETH",fig_1_1_9_data,"Ethiopia",50,1500,2);
  createFig_1_1_9_10("#fig_1_1_9_LSO",fig_1_1_9_data,"Lesotho",100,2500,2);
  createFig_1_1_9_10("#fig_1_1_9_NAM",fig_1_1_9_data,"Namibia",200,3000,2);
  createFig_1_1_9_10("#fig_1_1_9_SLE",fig_1_1_9_data,"Sierra Leone",50,600,2);
  createFig_1_1_9_10("#fig_1_1_9_VNM",fig_1_1_9_data,"Viet Nam",50,800,2);
  
  createFig_1_1_9_10("#fig_1_1_9_CHN",fig_1_1_9_data,"China",20,200,2);
  createFig_1_1_9_10("#fig_1_1_9_GAB",fig_1_1_9_data,"Gabon",100,1000,2);
  createFig_1_1_9_10("#fig_1_1_9_LBR",fig_1_1_9_data,"Liberia",50,600,2);
  createFig_1_1_9_10("#fig_1_1_9_NGA",fig_1_1_9_data,"Nigeria",20,400,2);
  createFig_1_1_9_10("#fig_1_1_9_ZAF",fig_1_1_9_data,"South Africa",200,3000,2);
  createFig_1_1_9_10("#fig_1_1_9_ZMB",fig_1_1_9_data,"Zambia",100,1500,2);
  
  createFig_1_1_9_10("#fig_1_1_10_KHM",fig_1_1_10_data,"Cambodia",100,1500,2);
  createFig_1_1_9_10("#fig_1_1_10_RUS",fig_1_1_10_data,"Russian Federation",20,200,2);
  createFig_1_1_9_10("#fig_1_1_10_ZWE",fig_1_1_10_data,"Zimbabwe",100,1500,2);

});  

```



